---
title: "Package 'TVLT'"
author: "Yizhen Xu, Tao Liu, Joseph Hogan"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 3
vignette: >
  %\VignetteIndexEntry{Targeted Viral Load Testing}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

**Type** Package

**Title** Targeted Viral Load Testing

**Version** 1.0

**Date** "`r Sys.Date()`"

**Authors** Yizhen Xu, Tao Liu, Joseph Hogan

**Maintainer** Yizhen (yizhen_xu@brown.edu)

**Description** This package implements the optimal allocation of viral load testing under constrained availability.

**License** ???

**URL**  https://github.com/yizhenxu/TVLT

**Depends** R (>= 3.2.0)

**LazyData** true

<!-- START doctoc -->
<!-- END doctoc -->

###Check.exp.tilt

*Check exponential tilt model assumption*

####Description

> This function provides graphical assessment to the suitability of the exponential tilt model for risk score in finding optimal tripartite rules by semiparametric approach.
$$g_1(s) = exp(\beta_0^*+\beta_1*s)*g_0(s)$$

####Usage

> Check.exp.tilt( Z, S)


####Arguments

> - **Z** &nbsp;&nbsp;&nbsp;&nbsp; True disease status (No disease / treatment success coded as Z=0, diseased / treatment failure coded as Z=1).
> - **S**  &nbsp;&nbsp;&nbsp;&nbsp;  Risk score.

####Value

> Returns the plot of empirical density for risk score S, joint empirical density for (S,Z=1) and (S,Z=0), and the density under the exponential tilt model assumption for (S,Z=1) and (S,Z=0).

####Author(s)

> Yizhen Xu (yizhen_xu@brown.edu), Tao Liu, Joseph Hogan

####References

> T. Liu, J. Hogan, L. Wang, S. Zhang, R. Kantor (2013) *Journal of the American Statistical Association* Vol.108, No.504

####Examples

```{r,eval=FALSE}
d = Simdata
Z = d$Z # True Disease Status
S = d$S # Risk Score
Check.exp.tilt( Z, S)
```
_____

###CV.TVLT

*Cross Validation*

####Description

> This function allows you to compute the average of misdiagnoses rate for viral failure and the optimal risk under min $\lambda$ rules from K-fold cross-validation.

####Usage

> CV.TVLT(Z, S, phi, K = 10, method = "semipar", lambda)


####Arguments

> - **Z** &nbsp;&nbsp;&nbsp;&nbsp; True disease status (No disease / treatment success coded as Z=0, diseased / treatment failure coded as Z=1).
> - **S**  &nbsp;&nbsp;&nbsp;&nbsp;  Risk score.
> - **phi**  &nbsp;&nbsp;&nbsp;&nbsp;  Percentage of patients taking viral load test.
> - **K**  &nbsp;&nbsp;&nbsp;&nbsp; Number of folds in cross validation. The default is 10.
> - **method**  &nbsp;&nbsp;&nbsp;&nbsp; The method to be used. The default is "nonpar", which returns result using nonparametric method. Another possible value is "semipar", which returns result estimated by semiparametric method assuming exponential tilt.
> - **lambda** &nbsp;&nbsp;&nbsp;&nbsp; A user-specified weight that reflects relative loss for the two types of misdiagnoses, taking value in [0,1]. $Loss=\lambda*I(FN)+(1-\lambda)*I(FP)$.
 
####Value

> Cross validated results on false classification rates (FNR, FPR), $\lambda-$ risk, total misclassification rate and AUC.

####Author(s)

> Yizhen Xu (yizhen_xu@brown.edu), Tao Liu, Joseph Hogan

####References

> T. Liu, J. Hogan, L. Wang, S. Zhang, R. Kantor (2013) *Journal of the American Statistical Association* Vol.108, No.504

####Examples

```{r,eval=FALSE}
d = Simdata
Z = d$Z # True Disease Status
S = d$S # Risk Score
phi = 0.1 #10\% of patients taking viral load test
lambda =0.5
CV.TVLT(Z, S, phi, K = 10, method = "semipar", lambda)
```


_____

###Opt.nonpar.rule

*Optimal Nonparametric Rule*

####Description

> This function gives you the optimal nonparametric tripartite rule that minimizes the min $\lambda$ rules.

####Usage

> Opt.nonpar.rule( Z, S, phi, lambda)


####Arguments

> - **Z** &nbsp;&nbsp;&nbsp;&nbsp; True disease status (No disease / treatment success coded as Z=0, diseased / treatment failure coded as Z=1).
> - **S**  &nbsp;&nbsp;&nbsp;&nbsp;  Risk score.
> - **phi**  &nbsp;&nbsp;&nbsp;&nbsp;  Percentage of patients taking viral load test.
> - **lambda** &nbsp;&nbsp;&nbsp;&nbsp; A user-specified weight that reflects relative loss for the two types of misdiagnoses, taking value in [0,1]. $Loss=\lambda*I(FN)+(1-\lambda)*I(FP)$.
 
####Value

> Optimal nonparametric rule and its associated misclassification rates (FNR, FPR), optimal lambda risk, and total misclassification rate (TMR). 

####Author(s)

> Yizhen Xu (yizhen_xu@brown.edu), Tao Liu, Joseph Hogan

####References

> T. Liu, J. Hogan, L. Wang, S. Zhang, R. Kantor (2013) *Journal of the American Statistical Association* Vol.108, No.504

####Examples

```{r,eval=FALSE}
d = Simdata
Z = d$Z # True Disease Status
S = d$S # Risk Score
phi = 0.1 #10\% of patients taking viral load test
lambda = 0.5
Opt.nonpar.rule( Z, S, phi, lambda)
```

_____

###Opt.semipar.rule

*Optimal Semiparametric Rule*

####Description

> This function gives you the optimal semiparametric tripartite rule that minimizes the min $\lambda$ rules.

####Usage

> Opt.semipar.rule( Z, S, phi, lambda)


####Arguments

> - **Z** &nbsp;&nbsp;&nbsp;&nbsp; True disease status (No disease / treatment success coded as Z=0, diseased / treatment failure coded as Z=1).
> - **S**  &nbsp;&nbsp;&nbsp;&nbsp;  Risk score.
> - **phi**  &nbsp;&nbsp;&nbsp;&nbsp;  Percentage of patients taking viral load test.
> - **lambda** &nbsp;&nbsp;&nbsp;&nbsp; A user-specified weight that reflects relative loss for the two types of misdiagnoses, taking value in [0,1]. $Loss=\lambda*I(FN)+(1-\lambda)*I(FP)$.
 
####Value

> Optimal semiparametric rule and its associated misclassification rates (FNR, FPR), optimal lambda risk, and total misclassification rate (TMR). 

####Author(s)

> Yizhen Xu (yizhen_xu@brown.edu), Tao Liu, Joseph Hogan

####References

> T. Liu, J. Hogan, L. Wang, S. Zhang, R. Kantor (2013) *Journal of the American Statistical Association* Vol.108, No.504

####Examples

```{r,eval=FALSE}
d = Simdata
Z = d$Z # True Disease Status
S = d$S # Risk Score
phi = 0.1 #10% of patients taking viral load test
lambda = 0.5
Opt.semipar.rule( Z, S, phi, lambda)
```
_____

###ROC.nonpar

*Nonparametric ROC Analysis*

####Description

> This function performs ROC analysis for tripartite rules by nonparametric approach. If 'plot=TRUE', the ROC curve is returned.

####Usage

> ROC.nonpar( Z, S, phi,plot=TRUE)

####Arguments

> - **Z** &nbsp;&nbsp;&nbsp;&nbsp; True disease status (No disease / treatment success coded as Z=0, diseased / treatment failure coded as Z=1).
> - **S**  &nbsp;&nbsp;&nbsp;&nbsp;  Risk score.
> - **phi**  &nbsp;&nbsp;&nbsp;&nbsp;  Percentage of patients taking viral load test.
 
####Value

> - **AUC**   &nbsp;&nbsp;&nbsp;&nbsp; The area under the ROC curve.
> - **FNR**   &nbsp;&nbsp;&nbsp;&nbsp; Misdiagnoses rate for viral failure (false negative rate).
> - **FPR**  &nbsp;&nbsp;&nbsp;&nbsp; Misdiagnoses rate for treatment failure (false positive rate).
 
####Author(s)

Yizhen Xu (yizhen_xu@brown.edu), Tao Liu, Joseph Hogan

####References

> T. Liu, J. Hogan, L. Wang, S. Zhang, R. Kantor (2013) *Journal of the American Statistical Association* Vol.108, No.504

####Examples

```{r,eval=FALSE}
d = Simdata
Z = d$Z # True Disease Status
S = d$S # Risk Score
phi = 0.1 #10% of patients taking viral load test
a = ROC.nonpar( Z, S, phi,plot=TRUE)
a$AUC
a$FNR
a$FPR
```
_____

###ROC.semipar

*Semiparametric ROC Analysis*

####Description 

> This function performs ROC analysis for tripartite rules by semiparametric approach. If 'plot=TRUE', the ROC curve is returned.

####Usage

> ROC.semipar( Z, S, phi,plot=TRUE)

####Arguments

> - **Z** &nbsp;&nbsp;&nbsp;&nbsp; True disease status (No disease / treatment success coded as Z=0, diseased / treatment failure coded as Z=1).
> - **S**  &nbsp;&nbsp;&nbsp;&nbsp;  Risk score.
> - **phi**  &nbsp;&nbsp;&nbsp;&nbsp;  Percentage of patients taking viral load test.
 
####Value

> - **AUC**   &nbsp;&nbsp;&nbsp;&nbsp; The area under the ROC curve.
> - **FNR**   &nbsp;&nbsp;&nbsp;&nbsp; Misdiagnoses rate for viral failure (false negative rate).
> - **FPR**  &nbsp;&nbsp;&nbsp;&nbsp; Misdiagnoses rate for treatment failure (false positive rate).
 
####Author(s)

Yizhen Xu (yizhen_xu@brown.edu), Tao Liu, Joseph Hogan

####References

> T. Liu, J. Hogan, L. Wang, S. Zhang, R. Kantor (2013) *Journal of the American Statistical Association* Vol.108, No.504

####Examples

```{r,eval=FALSE}
d = Simdata
Z = d$Z # True Disease Status
S = d$S # Risk Score
phi = 0.1 #10% of patients taking viral load test
a = ROC.semipar( Z, S, phi,plot=TRUE)
a$AUC
a$FNR
a$FPR
```

_____

###Rules.set

*Nonparametric Rules Set*

####Description

> This function gives you all possible cutoffs [l,u] for tripartite rules, by applying nonparametric search to the given data.
$$P(S \in [l,u]) \le \phi$$

####Usage

> Rules.set( Z, S, phi)

####Arguments

> - **Z** &nbsp;&nbsp;&nbsp;&nbsp; True disease status (No disease / treatment success coded as Z=0, diseased / treatment failure coded as Z=1).
> - **S**  &nbsp;&nbsp;&nbsp;&nbsp;  Risk score.
> - **phi**  &nbsp;&nbsp;&nbsp;&nbsp;  Percentage of patients taking viral load test.
 
####Value

> Matrix with 4 columns. Each row is a possible tripartite rule, with output on lower cutoff, upper cutoff and corresponding misclassification rates (FNR, FPR).
 
####Author(s)

> Yizhen Xu (yizhen_xu@brown.edu), Tao Liu, Joseph Hogan

####References

> T. Liu, J. Hogan, L. Wang, S. Zhang, R. Kantor (2013) *Journal of the American Statistical Association* Vol.108, No.504

####Examples

```{r,eval=FALSE}
d = Simdata
Z = d$Z # True Disease Status
S = d$S # Risk Score
phi = 0.1 #10\% of patients taking viral load test
Rules.set( Z, S, phi)
```

_____

###Semi.par.rule

*Min TMR Semiparametric Rule*

####Description

> This function gives you the optimal semiparametric tripartite rule that minimizes TMR (total misclassification risk).

####Usage

> Semi.par.rule( Z, S, phi)

####Arguments

> - **Z** &nbsp;&nbsp;&nbsp;&nbsp; True disease status (No disease / treatment success coded as Z=0, diseased / treatment failure coded as Z=1).
> - **S**  &nbsp;&nbsp;&nbsp;&nbsp;  Risk score.
> - **phi**  &nbsp;&nbsp;&nbsp;&nbsp;  Percentage of patients taking viral load test.
 
####Value

> Optimal semiparametric rule (lower and upper cutoff) and its associated minimized TMR.
 
####Author(s)

> Yizhen Xu (yizhen_xu@brown.edu), Tao Liu, Joseph Hogan

####References

> T. Liu, J. Hogan, L. Wang, S. Zhang, R. Kantor (2013) *Journal of the American Statistical Association* Vol.108, No.504

####Examples

```{r,eval=FALSE}
d = Simdata
Z = d$Z # True Disease Status
S = d$S # Risk Score
phi = 0.1 #10\% of patients taking viral load test
Semi.par.rule( Z, S, phi)
```
_____

###Simdata

*Simulated data for package illustration*

####Description

> A simulated dataset containing true disease status, CD4 count, and risk score. Simulation parameters are determined by fitting a dataset from  Miriam Immunology Clinic. See details.

####Format

> A data frame with 5000 simulated observations on the following 3 variables.
> - **Z** &nbsp;&nbsp;&nbsp;&nbsp; True disease status (No disease / treatment success coded as Z=0, diseased / treatment failure coded as Z=1).
> - **CD4**  &nbsp;&nbsp;&nbsp;&nbsp;  Simulated CD4 count with different distribution with respect to different disease status.
> - **S**  &nbsp;&nbsp;&nbsp;&nbsp;  Risk score defined as the negative value of CD4 count.
 
####Details

> We first simulate viral status $Z$ assuming $Z\sim Bernoulli(p)$ with $p=0.25$ which was estimated from the Miriam Immunology Clinic data; and then conditonal on $Z$, simulate ${CD4|Z=z}=ceiling(W)$ with $W\sim Gamma(\eta_z,\kappa_z)$ where $\eta$ and $\kappa$ are shape and scale parameters.$(\eta_0,\kappa_0)=(3.2,152)$ and $(\eta_1,\kappa_1)=(2.3,133)$. Parameter values are chosen as maximum likelihood estimates obtained by fitting gamma distributions to the Miriam Immunology Clinic data.
 
####Author(s)

> Yizhen Xu (yizhen_xu@brown.edu), Tao Liu, Joseph Hogan

####References

> T. Liu, J. Hogan, L. Wang, S. Zhang, R. Kantor (2013) *Journal of the American Statistical Association* Vol.108, No.504

####Examples

```{r,eval=FALSE}
data(Simdata)
summary(Simdata)
plot(Simdata)
```
_____

###TVLT.summ

*Summary of Disease Status and Risk Score*

####Description

> This function allows you to compute the percentage of diseased (equivalent to treatment failure Z=1) and show distribution summary of risk score (S) by the true disease status (Z).

####Usage

> TVLT.summ(Z,S)

####Arguments

> - **Z** &nbsp;&nbsp;&nbsp;&nbsp; True disease status (No disease / treatment success coded as Z=0, diseased / treatment failure coded as Z=1).
> - **S**  &nbsp;&nbsp;&nbsp;&nbsp;  Risk score.


####Value

> Percentage of treatment failure; Summary statistics (mean, standard deviation, minimum, median, maximum and IQR) of risk score by true disease status; Distribution plot.
 
####Author(s)

> Yizhen Xu (yizhen_xu@brown.edu), Tao Liu, Joseph Hogan

####References

> T. Liu, J. Hogan, L. Wang, S. Zhang, R. Kantor (2013) *Journal of the American Statistical Association* Vol.108, No.504

####Examples

```{r,eval=FALSE}
d = Simdata
Z = d$Z # True Disease Status
S = d$S # Risk Score
TVLT.summ(Z,S)
```